A thermodynamic view of dusty protoplanetary disks: axisymmetric stability

### 2016-11-02 linear.tex
regular update
work/home sync

However, retaining all the terms (but still in the incompressible
limit) can yield instability, as follows. Eq. \ref{kz_zero_si} admit
pure oscillations with $\sigma = -\omega$ when 
\begin{align*}
  \omega &= \pm \kappa, \\
  \tstop & = \frac{\omega}{ k _x F_r }
         \equiv t_{\mathrm{s}0}.
\end{align*} 
Since $\tstop\geq 0$, we require $k_xF_r > 0$ if $\omega = \kappa$;
and $k_xF_r < 0$ if $\omega = -\kappa$. For $k_x>0$ these marginally
stable epicycles require radial pressure gradients decreasing and increasing
outwards, respectively.

@@

We now perturb about this state of marginal stability by writing
$\sigma = \dd \sigma - \omega $ with $\dd\sigma = \ii\dd s -
\dd\omega$, $\tstop = t_{\mathrm{s}0} +
\dd\tstop $, and assume the $\dd$ quantities are small in
magnitude. Then 
\begin{align*} 
\dd s = \frac{2\epsilon \kappa^2 }{1 +
  4\epsilon^2 t_{\mathrm{s}0}^2\kappa^2}\dd \tstop. 
\end{align*}
Thus given $k_xF_r$, there exist growing modes 
when the stopping time is slightly larger than
$t_{\mathrm{s}0}$.

%{\bf note: several approx used}

If we instead fix $\tstop$ and vary $k_xF_r$ about marginal stability,
a similar exercise yields  
\begin{align*} 
  \dd s = \frac{\epsilon \tstop^3}{1 +
  4\epsilon^2 \tstop^2 \kappa^2}\dd \left(k_x^2F_r^2\right).  
\end{align*}
Thus it is possible to destabilize the system by slightly increasing
$|k_xF_r|>\kappa/ \tstop$ (i.e. shorter radial wavelengths and/or
stronger pressure gradients).

@@

To demonstrate the above instabilities we numerically solve the
full dispersion relation, Eq. \ref{streaming_dispersion} at fixed $K_x
= 1500$ and $\tepsilon=2$. Growth rates are shown in 
Fig. \ref{dusty_growth} as a function of $K_z$ at fixed $\taus\OmK=0.01$
(top panel, which includes `linC' in Table \ref{si_compare}); and as a
function of $\taus\OmK$ at fixed $K_z=50$ (bottom panel).

@@

>>>>

Now, writing $\sigma = \ii s - \omega$, the real and imaginary parts of 
Eq. \ref{kz_zero_si}  are
\begin{align}
\omega^2 - s^2 - \frac{s}{\epsilon\tstop}  - \kappa^2 &= 0, \label{epi_overs1}\\
\omega\left(2 s + \frac{1}{\epsilon\tstop}\right) - \frac{k_xF_r}{\epsilon} & =
                                                                     0.  \label{epi_overs2}
\end{align}
The growth rate satisfies
\begin{align*}
\left( \frac{k_xF_r\tstop}{1 + 2 s\epsilon\tstop}\right)^2 - s^2
  - \frac{s}{\epsilon \tstop} = \kappa^2. %\omega = \frac{k_xF_r\tstop}{1 + 2 s\epsilon\tstop}
\end{align*}

Assuming slow growth such that $\epsilon s \tstop \ll 1$, we find 
\begin{align}
  s \simeq \epsilon \tstop \left(k_xF_r\tstop\right)^2 - \kappa^2,
\end{align}
Growth is possible for
sufficiently large stopping times, and/or large radial pressure
gradients, and/or sufficiently small radial wavelengths. 
%For small growth rates the real frequency $|\omega| =  
%O(\kappa)$, so these are overstable epicycles.

It is interesting to note that Eq. \ref{epi_overs1}---\ref{epi_overs2}
admit pure oscillations with $s \equiv 0$, $\omega = 
\pm\kappa$ if 
\begin{align}
  \tstop = \frac{\kappa}{k_x\left|F_r\right|}. 
\end{align}
This is the condition for marginal stability.

% However, retaining all the terms (but still in the incompressible
% limit) can yield instability, as follows. Eq. \ref{kz_zero_si} admit
% pure oscillations with $\sigma = -\omega$ when 
% \begin{align*}
%   \omega &= \pm \kappa, \\
%   \tstop & = \frac{\omega}{ k _x F_r }
%          \equiv t_{\mathrm{s}0}.
% \end{align*} 
% Since $\tstop\geq 0$, we require $k_xF_r > 0$ if $\omega = \kappa$;
% and $k_xF_r < 0$ if $\omega = -\kappa$. For $k_x>0$ these marginally
% stable epicycles require radial pressure gradients decreasing and increasing
% outwards, respectively.

@@

% We now perturb about this state of marginal stability by writing
% $\sigma = \dd \sigma - \omega $ with $\dd\sigma = \ii\dd s -
% \dd\omega$, $\tstop = t_{\mathrm{s}0} +
% \dd\tstop $, and assume the $\dd$ quantities are small in
% magnitude. Then 
% \begin{align*} 
% \dd s = \frac{2\epsilon \kappa^2 }{1 +
%   4\epsilon^2 t_{\mathrm{s}0}^2\kappa^2}\dd \tstop. 
% \end{align*}
% Thus given $k_xF_r$, there exist growing modes 
% when the stopping time is slightly larger than
% $t_{\mathrm{s}0}$.

% %{\bf note: several approx used}

% If we instead fix $\tstop$ and vary $k_xF_r$ about marginal stability,
% a similar exercise yields  
% \begin{align*} 
%   \dd s = \frac{\epsilon \tstop^3}{1 +
%   4\epsilon^2 \tstop^2 \kappa^2}\dd \left(k_x^2F_r^2\right).  
% \end{align*}
% Thus it is possible to destabilize the system by slightly increasing
% $|k_xF_r|>\kappa/ \tstop$ (i.e. shorter radial wavelengths and/or
% stronger pressure gradients).

@@

To demonstrate the above instabilities we numerically solve the
full dispersion relation, Eq. \ref{streaming_dispersion} at fixed $K_x
= 1500$ and $\epsilon=2$. Growth rates are shown in 
Fig. \ref{dusty_growth} as a function of $K_z$ at fixed $\taus\OmK=0.01$
(top panel, which includes `linC' in Table \ref{si_compare}); and as a
function of $\taus\OmK$ at fixed $K_z=50$ (bottom panel).

@@

\begin{align*}
  s = \epsilon \left(1 - \tepsilon
  \right)\taus\OmK^2\left[4K_x^2\left(1-\tepsilon\right)^4\left(\taus\OmK\right)^2
  -1 \right]
\end{align*}
